---
title: "Lab 2: Intro to Markdown"
author: "Geog 418/518"
header-includes:
    - \usepackage{setspace}\doublespacing
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
fontsize: 12pt
spacing: double
---

```{r setup, include=FALSE}
# Help with markdown in general (Can get very detailed and in the weeds)
# https://bookdown.org/yihui/rmarkdown-cookbook/

# RCheat Sheet
# https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf

# Help with markdown formulas
# https://rpruim.github.io/s341/S19/from-class/MathinRmd.html

#####
#Load Libraries
library("knitr")
library("terra")
library("lubridate")
library("e1071")
library("gridExtra")
library("gtable")
library("grid")
library("ggplot2")
library("dplyr")
library("bcmaps")
library("tmap")
library("sf")
library("nngeo")

#####
#Set working directory
dir <-"C:/Users/philk/Desktop/Spring 2022/Geog 418/Assignment1_v2/PROT_HISTORICAL_INCIDENTS_SP"
knitr::opts_knit$set(root.dir = dir)

##  "E:/GEOG_418_2022/A1/Wednesday" 

```

## Introduction

Over the past decade, western North America has experienced substantial adverse effects from climate change, including an increase in the intensity and duration of wildfires. In British Columbia (BC), Canada, the 2018 wildfire season burned over 1.3 million hectares of land, surpassing the 2017 wildfire season (1.2 million hectares) for the highest ever recorded area burned in BC’s history [1]. The cost of wildfire suppression in 2018 totalled $615 million, while the second longest ever provincial state of emergency was enacted in 2017 for 10 weeks in response to the fires (only surpassed by the state of emergency declared in 2020 for the COVID-19 pandemic) [1]. With climate change contributing to ever-increasing wildfire risk in BC [2], statistical analyses of both fire sizes and spatial distribution can aid in assessing potential economic, health, and environmental impacts. This report seeks to statistically analyze the 2020 BC wildfire season based on fire size and geographic location.


```{r Data Cleaning, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
#####
#Read in data and extract attribute table
shp <- vect("./H_FIRE_PNT_point.shp") #read in shp file from current (".") working directory


bc <- vect(bc_neighbours()) #Get shp of BC bounds
bc <- bc[which(bc$name == "British Columbia" ),] #Extract just the BC province

shp <- shp[bc]

df <- as.data.frame(shp) #Extract the attribute table as a dataframe, store in variable

#We will need to subset the data into an given year
df <- subset(df, df$FIRE_YEAR == 2020)

#We will also need to know the date and the month of each fire
df$IGN_DATE <- as.Date(df$IGN_DATE, "%Y%m%d") #Convert to date string
df$IGN_Day <- yday(df$IGN_DATE) #Make new column with day of year for the ignition date
df$IGN_Month <- month(df$IGN_DATE, label = TRUE, abbr = TRUE) #create new column with month of ignition

#Remove values with an NA date
df <- subset(df, !is.na(df$IGN_DATE))

#Calculate new year column from date and check range
df$IGN_Year <- year(df$IGN_DATE)

#####
#Clean data to appropriate year and calculate descriptive statistics
df_year <- df[which(df$IGN_Year == 2020),] #subset only your year


##*****
######
#Check for 0 fire size
minSize <- min(df_year$SIZE_HA)

df_year <- subset(df_year, df_year$SIZE_HA > 0)

#Check 0's were removed
minSize <- min(df_year$SIZE_HA)
##*****
```

## Study Area
As mentioned above, this study focuses on the distribution of 2020 fire sizes and locations in BC. A map of the study area and fire points are shown below for reference (Figure 1).

```{r Study Area Map, echo=FALSE, eval=TRUE, warning=FALSE, fig.cap="British Columbia (BC) fire incidents in 2020."}

coords <- cbind(df_year$LONGITUDE, df_year$LATITUDE) #Store coordinates in new object
crs <- "+init=epsg:4326" #store the coordinate system in a new object
firePoints <- vect(coords, atts = df_year, crs = crs)#Make new spatial Points object using coodinates, data, and projection


StudyArea <- tm_shape(sf::st_as_sf(bc)) + #make the main shape
  tm_polygons(col = "gray50", border.col = "black") +  #fill polygons
  tm_shape(sf::st_as_sf(firePoints)) +
  tm_symbols(size = 0.05, col = "blue", shape = 20, alpha = 0.25) +
  tm_layout(title = "BC Fire Incidents 2020", 
            title.position = c("LEFT", "BOTTOM"),
            legend.position = c("LEFT", "TOP"))

StudyArea
```

## Data

Data for this study were collected from the BC open data catalog, retrieved from [here](https://catalogue.data.gov.bc.ca/dataset/fire-incident-locations-historical). This dataset has contains historical fire observations from 1950 until 2021 and contains a variety of information about the fires such as the location at which it occurred, and the size of the fire [1].

## Analysis

### Mean

The mean of 2020 fire size was calculated using the following formula:


$$
\bar{X_i} = \frac{\sum_{i = 1}^{n}X_i}{n_i}
$$

### Standard deviation

The standard deviation of 2020 fire size was calculated using the following formula:

$$
s = \sqrt{\frac{\sum (X_i - \bar{X})^2}{n - 1}}
$$

### Coefficient of Variation
The coefficient of variation of 2020 fire size was calculated using the following formula:

$$
CV = \frac{s}{\bar{X}}(100)
$$

## Results

The fire data points for BC’s 2020 fire season are categorized into summer fire points and total (population) fire points. The summer points encompass the data between July 1st and August 31st, while the total points cover data from the full year. Table 1 showcases a relatively small median and even smaller mode of fire sizes for both the summer and total points (each well under 1 hectares). However, the mean fire sizes are over 20 and 30 hectares for the total and summer points, respectively, while the standard deviation for both indicates a very large range in the data. Table 2 indicates that the data is positively skewed with a leptokurtic distribution (ie. the majority of fires sizes are very similar). The outliers contributing to this positive skewness can be seen in Figures 2, with one fire indicated as covering almost 8000 hectares. In terms of monthly differences, August had the highest total fire coverage, while the lowest was in June (Figure 3). The geographic distribution of the total fire points (Figures 4) showcases that the majority of fires occurred in southern BC, while the mean centre is further north but still within the southern half of the province.


```{r Descriptive Stats, echo=FALSE, eval=TRUE, warning=FALSE}
#Mean
meanPop <- mean(df_year$SIZE_HA) #This could produce a wrong value (NA) due to a single NA value in data
meanPop <- mean(df_year$SIZE_HA, na.rm = TRUE) #Use na.rm = TRUE to ignore NA values in calculation

#Check this website for the correct day numbers of your year: https://miniwebtool.com/day-of-the-year-calculator/
meanSummer <- mean(subset(df_year, IGN_Day >= 182 & IGN_Day <= 243)$SIZE_HA) #Calculate the mean fire size between July 1 and Aug 31

#Standard Deviation
sdPop <- sd(df_year$SIZE_HA, na.rm = TRUE) #Calculate the SD, ignoring NA values
sdSummer <- sd(subset(df_year, IGN_Day >= 182 & IGN_Day <= 243)$SIZE_HA) #Calculate the SD, ignoring NA values only for the summer months

#Mode
modePop <- as.numeric(names(sort(table(df_year$SIZE_HA), decreasing = TRUE))[1]) #make frequency table of fire size variable and sort it in desending order and extract the first row (Most Frequent)

#It will be cleaner if we use a new variable for the summer data
df_Summer <- subset(df_year, IGN_Day >= 182 & IGN_Day <= 243) #Make new variable to hold summer data
#make frequency table of fire size variable and sort it in desending order and extract the first row (Most Frequent)
modeSummer <- as.numeric(names(sort(table(df_Summer$SIZE_HA), decreasing = TRUE))[1])

#Median
medPop <- median(df_year$SIZE_HA, na.rm = TRUE)
medSummer <- median(df_Summer$SIZE_HA, na.rm = TRUE)

#Skewness
skewPop <- skewness(df_year$SIZE_HA, na.rm = TRUE)[[1]]
skewSummer <- skewness(df_Summer$SIZE_HA, na.rm = TRUE)[1]

#Kurtosis
kurtPop <- kurtosis(df_year$SIZE_HA, na.rm = TRUE)[1]
kurtSummer <- kurtosis(df_Summer$SIZE_HA, na.rm = TRUE)[1]

#CoV
CoVPop <- (sdPop / meanPop) * 100
CoVSummer <- (sdSummer / meanSummer) * 100

#Normal distribution test
normPop_PVAL <- shapiro.test(df_year$SIZE_HA)$p.value
normSummer_PVAL <- shapiro.test(df_Summer$SIZE_HA)$p.value


#####
#Create a table of descriptive stats
samples = c("Population", "Summer") #Create an object for the labels
means = c(meanPop, meanSummer) #Create an object for the means
sd = c(sdPop, sdSummer) #Create an object for the standard deviations
median = c(medPop, medSummer) #Create an object for the medians
mode <- c(modePop, modeSummer) #Create an object for the modes
skewness <- c(skewPop, skewSummer) #Create an object for the skewness
kurtosis <- c(kurtPop, kurtSummer) #Create an object for the kurtosis
CoV <- c(CoVPop, CoVSummer) #Create an object for the CoV
normality <- c(normPop_PVAL, normSummer_PVAL) #Create an object for the normality PVALUE


means <- round(means, 3)
sd <- round(sd, 3)
median <- round(median, 3)
mode <- round(mode, 3)
skewness <- round(skewness, 3)
kurtosis <- round(kurtosis, 3)
CoV <- round(CoV, 3)
normality <- signif(normality, digits = 3)

data.for.table1 = data.frame(samples, means, sd, median, mode)
data.for.table2 = data.frame(samples, skewness, kurtosis, CoV, normality)

kable(data.for.table1, caption = paste0("Descriptive statistics for ", 2020, " fire season"))
kable(data.for.table2, caption = paste0("Dispersion statistics for ", 2020, " fire season"))
```



```{r Histogram, echo=FALSE, eval=TRUE, warning=FALSE, fig.cap="Frequency of fire sizes for the year 2020."}
#Set limits for Y axis dynamically
histogram <- ggplot(df_year, aes(x = SIZE_HA)) + #Create new GGplot object with data attached and fire size mapped to X axis
  geom_histogram(bins = 50, color = "black", fill = "white") + #make histogram with 30 bins, black outline, white fill
  labs(title = "", x = "Size of Wild Fire (ha)", y = "Frequency") + #label plot, x axis, y axis
  theme_classic() + #set the theme to classic (removes background and borders etc.)
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) + #set title to center and bold
  scale_y_continuous(breaks = seq(0, 800, by = 100)) # set y axis labels to 0 - 700 incrimenting by 100
histogram
```


```{r Barchart, echo=FALSE, eval=TRUE, warning=FALSE, fig.cap="Total Burned Area by Month for the year 2020."}
#Set limits for Y axis dynamically
#Total Size by Month GGPLOT
barGraph <- df_year %>% #store graph in bargraph variable and pass data frame as first argument in next line
  group_by(IGN_Month) %>% #use data frame and group by month and pass to first argument in next line
  summarise(sumSize = sum(SIZE_HA, na.rm = TRUE)) %>% #sum up the total fire size for each month and pass to GGplot
  ggplot(aes(x = IGN_Month, y = sumSize)) + #make new GGPLOT with summary as data and month and total fire size as x and y
  geom_bar(stat = "identity") + #make bar chart with the Y values from the data (identity)
  labs(title = "", x = "Month", y = "Total Burned Area (ha)") + #label plot, x axis, y axis
  theme_classic() + #set the theme to classic (removes background and borders etc.)
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) +
  scale_y_continuous(breaks = seq(0, 14000, by = 2000))
barGraph
```


```{r MeanCenter, echo=FALSE, eval=TRUE, warning=FALSE, fig.cap="Fire locations, mean center, and standard ellipse for the year 2020."}
#####
#Creating maps
bc <- vect(bc_neighbours()) #Get shp of BC bounds
bc <- bc[which(bc$name == "British Columbia" ),] #Extract just the BC province

#####
#Making Maps with tm package
#Make spatial object out of data
coords <- cbind(df_year$LONGITUDE, df_year$LATITUDE) #Store coordinates in new object
crs <- "+init=epsg:4326" #store the coordinate system (CRS) in a new object

firePoints <- vect(coords, atts = df_year, crs = crs) #Make new spatial Points object using coodinates, data, and projection

meanCenter <- data.frame(name = "Mean Center of fire points", long = mean(df_year$LONGITUDE), lat = mean(df_year$LATITUDE))

coords2 <- cbind(meanCenter$long, meanCenter$lat)
crs2 <- "+init=epsg:4326"
meanCenterPoint <- vect(coords2, atts = meanCenter, crs = crs2)

SDX <- sd(df_year$LONGITUDE)
SDY <- sd(df_year$LATITUDE)

ellipse <- st_ellipse(sf::st_as_sf(meanCenterPoint), SDX, SDY, res = 60)

map_TM <- tm_shape(sf::st_as_sf(bc)) + 
  tm_fill(col = "gray50") +  
  tm_shape(sf::st_as_sf(firePoints)) +
  tm_symbols(col = "red", alpha = 0.1) +
  tm_shape(sf::st_as_sf(meanCenterPoint)) +
  tm_symbols(col = "blue", alpha = 0.8) +
  tm_shape(sf::st_as_sf(ellipse)) + 
  tm_fill(col = "lightblue", alpha = 0.3) +
  tm_credits(paste("Mean Center: ", round(meanCenter$long, 3), ", ", round(meanCenter$lat, 3), sep = ""), position = c("RIGHT", "BOTTOM")) +
  tm_layout(title = paste0("BC Fire Locations ", 2020), title.position = c("LEFT", "BOTTOM"))
map_TM
```

## Conclusions

Overall, the fire size data indicates a very skewed distribution for both the total and
summer coverage, albeit the summer fire sizes are somewhat less skewed than the total. These
results indicate that although the vast majority of fires were very small, several drastically larger fires were recorded over the course of the season, highlighting the complexity in predicting both fire scope and intensity for any given season. However, analysis of these results also reveals important patterns in the data, such as fire timing and distribution. For example, even though the fire season covers a six month period, the majority of fires occurred over one month. Additionally, the data reveals a southerly geographic distribution pattern. While these analyses are useful when singularly looking at the 2020 fire year, their applicability in a broader context cannot be assessed without comparison between previous fire seasons. Through such
comparisons however, seasonally derived statistical results can help point to larger fire trends,
thus helping with predictability modelling and subsequent enaction of appropriate response
measures.


## References

1. Province of British Columbia. (2020). Wildfire season summary. Retrieved from the Government
of British Columbia website: https://www2.gov.bc.ca/gov/content/safety/wildfirestatus/about-bcws/wildfire-history/wildfire-season-summary.

2. Kirchmeier‐Young, M. C., Gillett, N. P., Zwiers, F. W., Cannon, A. J., & Anslow, F. S. (2018).
Attribution of the influence of human‐induced climate change on an extreme fire season. Earth's Future, 7(1), 2-10.

